Estimating autocorrelation from multiple Markov chains

### 2020-06-24 Paper/main.tex
Cleaned the appendix!

In the next section we introduce the new sample autocovariance estimator and calculate its bias for finite samples. We observe that the bias term for R-ACF is similar to the bias of empirical ACF except for a few small order end-effect terms that vanish as $n \to \infty$. In section \ref{sec:variance_est}, we introduce three estimators for $\Sigma$ and study the properties of RSV estimator through proof of strong consistency, bias, and variance calculations. We will elaborate on the estimator for ESS using RSV in section \ref{sec:ess}. Section \ref{sec:examples} gives a simulation study on three different examples for the proposed estimators. There are only a handful of stochastic processes where the true autocovariance and $\Sigma$ are known. To compare the estimators when the truth is known, we use a slowly mixing vector autoregressive process of order 1 (VAR(1)). We succesfully show in section \ref{ex:var} that our estimators yield better results as compared to the classical sample ACF and SVE. A more promising advantage of our estimators is observed when the target distribution is multi-modal. For this purpose we use a bivariate bi-modal probability distribution introduced by \cite{gelman1991note} in section \ref{ex:boomerang}. All the important proofs are given in the Appendix.

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[h]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{scatter_plot_500.pdf}
         \caption{$n = 500$}
         \label{subfig:var-sp_500}
     \end{subfigure}
     \begin{subfigure}[h]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{scatter_plot_50000.pdf}
         \caption{$n = 5 \times 10^4$}
         \label{subfig:var-sp_50000}
     \end{subfigure}
    \caption{Scatter plot of five Markov chains for different sample sizes.}
    \label{fig:var-sp}
\end{figure}

@@

\begin{lemma} \label{lemma:consis_1}
  If Assumption~\ref{ass:sip} holds, then $\| \bar{\bar{X}} - \mu\|_{\infty}, \|\bar{X}_s - \mu\|_{\infty} \xrightarrow[]{a.s.} 0$ as $n \to \infty$ for all $s \in \{1,..., m\}$\\
  \end{lemma}

\begin{proof}
Let $\|.\|$ denote the Euclidean norm. Assumption \ref{ass:sip} allows us to upper bound $\|\bar{X}_s - \mu\|_{\infty}$ using SIP. Following that, we show that the upper bound term converges to 0 as $n \to \infty$ using lemma \ref{lemma: brownian}.
 \begin{align*}
    \|\bar{X}_s - \mu\|_{\infty} & \leq \|\bar{X}_s - \mu\| \\
    &= \dfrac{1}{n}\left\|\sum_{t=1}^{n}X_{st} - n\mu\right\| = \dfrac{1}{n}  \left \|\sum_{t=1}^{n}X_{st} - n\mu \pm L B(n) \right\|\\
    & \leq \dfrac{1}{n}\left\|\sum_{t=1}^{n}X_{st} - \mu - L B(n)\right\| + \dfrac{\left\|L B(n)\right\|}{n}\\
    &< \dfrac{D\psi(n)}{n} + \dfrac{\|L B(n)\|}{n}\\
    &< \dfrac{D\psi(n)}{n} + \dfrac{1}{n}\|L\| \left(\sum\limits_{i=1}^{p}|B^{(i)}(k)|^2\right)^{1/2}\\
    & \leq \dfrac{D\psi(n)}{n} + \dfrac{1}{n}\|L\| p^{1/2}(1+\epsilon)\sqrt{2n \log\log n}\\
    & \xrightarrow[]{a.s.} 0\;\; as \;\; n\to \infty
 \end{align*}
%
Similarly one can easily show that $\|\bar{\bar{X}} - \mu\|_{\infty} \to 0$ as $n \to \infty$ using the same upper-bound as:
% 
\begin{align*}
    \|\bar{\bar{X}} - \mu\|_{\infty} & \leq \|\bar{\bar{X}} - \mu\| = \dfrac{1}{m}\left \|\sum_{j=1}^{m}(\bar{X}_j- \mu) \right\| \leq \dfrac{1}{m}\sum_{j=1}^{m}\|\bar{X}_j - \mu\| \xrightarrow{a.s.} 0 \, \textrm{as } n \to \infty 
\end{align*}
\end{proof}

@@

\begin{proof}[Proof of Theorem~\ref{th:RAC_expec}]
\begin{align*}
    & \hat{\Gamma}_{RAC,s}(k) \\
   & = \dfrac{1}{n}\sum_{t=1}^{n-|k|} \left(X_{st} - \bar{\bar{X}} \right) \left(X_{s(t+k)} - \bar{\bar{X}} \right)^T \\
    &= \left[\dfrac{1}{n} \sum_{t=1}^{n-|k|} \left(X_{st} - \bar{X}_s \right) \left(X_{s(t+k)} - \bar{X}_s \right)^T \right] + \left[\dfrac{1}{n} \sum_{t=1}^{|k|} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - X_{st} \right)^T\right] \\ 
    & \quad +  \left[\dfrac{1}{n} \sum_{t=n-|k|+1}^{n}  \left( \bar{X}_s - X_{st} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right] + \left[\dfrac{n-|k|}{n} \left(\bar{X}_s - \bar{\bar{X}}\right)   \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \hat{\Gamma}_s(k) - \dfrac{1}{n} \sum_{t=1}^{|k|}A_{st}^T - \dfrac{1}{n}\sum_{t=n-|k|+1}^{n}A_{st} + \dfrac{n-|k|}{n}  \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,, \numberthis \label{eq:acf_breakdown}
\end{align*}
%
where $A_{st} = (X_{st}-\bar{X}_s)(\bar{X}_s - \bar{\bar{X}})^T$. We will study the expectations of the each of the above terms. Without loss of generality, consider $A_{11}$,
\begin{align*}
    &\E \left[A_{11} \right]\\
     &= \mathbb{E} \left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]\\
     &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E}\left[X_{11}\bar{\bar{X}}^T \right] + \mathbb{E} \left[\bar{X}_1\bar{\bar{X}}^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{1}{m}\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{m-1}{m}\mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \dfrac{1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right] + \dfrac{m-1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \dfrac{m-1}{m}\left(\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{t=1}^{n}\mathbb{E} \left[X_{11}X_{1t}^T \right] - \mathbb{E}\left[X_{11}\right] \mathbb{E}\left[\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\right]\mathbb{E}\left[\bar{X}_2^T\right] - \text{Var}\left[\bar{X}_1 \right] - \mathbb{E}\left[\bar{X}_1 \right]\mathbb{E}\left[\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{k=0}^{n-1}\Gamma(k) + \mu \mu^T - \mu \mu^T + \mu \mu^T - \text{Var}\left[\bar{X}_1 \right] - \mu \mu^T\right)\\
    &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right]\right)\,. \numberthis \label{eq:acf_2}
\end{align*}
%
Similarly,
%    
\begin{align}
\label{eq:acf_3}
    \mathbb{E} \left[ A_{11}^T \right] &= \mathbb{E}\left[ A_{11}\right]^T = \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right] \right)^T\,.
    % &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)\\
\end{align}

Further,
\begin{align*}
\mathbb{E}\left[ \left(\bar{X}_{s} - \bar{\bar{X}} \right)  \left(\bar{X}_{s} - \bar{\bar{X}} \right)^{T}\right] &= \mathbb{E}\left[ \bar{X}_{s}\bar{X}_{s}^{T} - \bar{X}_{s}\bar{\bar{X}}^{T} - \bar{\bar{X}} \bar{X}_{s}^{T} + \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= \mathbb{E}\left[ \bar{X}_{s}\bar{X}_{s}^{T} - \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= \left(\mathbb{E}\left[\bar{X}_{1}\bar{X}_{1}^{T}\right] - \mathbb{E}\left[\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\right)\\
&= \left(\Var(\bar{X}_{1}) + \mu\mu^{T} - \Var(\bar{\bar{X}}) - \mu\mu^{T}\right)\\
&= \dfrac{m-1}{m}\Var(\bar{X}_1)\,. \numberthis \label{eq:acf_4}
\end{align*}
%
Using \eqref{eq:priestly}, \eqref{eq:acf_2}, \eqref{eq:acf_3}, and \eqref{eq:acf_4} in \eqref{eq:acf_breakdown},
\begin{align*}
& \E \left[\hat{\Gamma}_{RAC,s}(k) \right] \\
&= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - \dfrac{1}{n} \left(\sum\limits_{t=1}^{|k|}\mathbb{E}[A_{1t}^T] + \sum\limits_{t=n-|k|+1}^{n}\mathbb{E}[A_{1t}]\right) + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - {\color{red}\dfrac{2|k|}{n}\dfrac{m-1}{mn}\left(\sum_{h=0}^{n-1}\Gamma(h) - n \Var(\bar{X}_1)\right)} + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\,.
\end{align*}
{\color{red} DV: Instead of the ``2'', should it be something that has $\Gamma(h) + \Gamma(h)^T$?}

% Using the results of \cite{song1995optimal} given in \eqref{eq:song} to expand $\Var(\bar{X}_1)$,
% \[
% \mathbb{E}\left[\hat{\Gamma}_{RAC}(k)\right] = \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + \dfrac{2(m-1)}{m}\dfrac{\abs{k}}{n^2} \left(\Sigma - \sum_{h=0}^{n-1}\Gamma(h)\right) + o(n^{-2})\,.
% \]
%
% We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.

For a diagonal element of $\Gamma$,
\begin{align*}
& \E \left[ \hat{\Gamma}_{RAC,s}^{ii}\right]\\
& = \mathbb{E}\left[\hat{\Gamma}_{s}^{ii}(k)\right] - {\color{red}\dfrac{2|k|}{n}\dfrac{m-1}{mn}\left(\sum_{h=0}^{n-1}\Gamma^{ii}(h) - n \Var(\bar{X}_1)^{ii}\right)} + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)^{ii}\,.
\end{align*}
In the presence of positive correlation, the leftover term is positive. 
\end{proof}

@@

Consequently
\[
\left|\hat{\Sigma}_{RSV}^{ij} - \tilde{\Sigma}^{ij}  \right| = |M_1 + M_2| \leq |M_1| + |M_2|\,.
\]
Consider,
\begin{align*}
& |M_1| \\ 
   & = \left|\dfrac{1}{m}\sum _{s=1}^{m}\left\{\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\sum_{t=1}^{n-|k|} \dfrac{1}{n}\left[ \left(X_{st} - \mu \right)_i  \left(\mu-\bar{\bar{X}} \right)_j + \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)} - \mu \right)_j\right]\right\}\right| \\
   & \leq \dfrac{1}{m}\sum_{s=1}^{m}\left\{\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\left[ \dfrac{1}{n}{\left|\sum\limits_{t=1}^{n-|k|} \left(X_{st} - \mu \right)_i  \left(\mu - \bar{\bar{X}} \right)_j\right|} + \dfrac{1}{n}{\left|\sum_{t=1}^{n-|k|} \left(\mu-\bar{\bar{X}} \right)_i\left(X_{s(t+k)} - \mu \right)_j\right|} \right]\right\}\\
    & \leq \dfrac{1}{m}\sum_{s=1}^{m}\left\{\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\left[ \dfrac{1}{n}\left|\sum_{t=1}^{n-|k|}(X_{st}- \mu)_i\right|\left|(\mu-\bar{\bar{X}})_j\right|+ \dfrac{1}{n}\left|(\mu-\bar{\bar{X}})_i\right|\left|\sum_{t=1}^{n-|k|}(X_{j(t+k)}-\mu)_j\right|\right]\right\}\\
    & \leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \sum_{s=1}^{m}\sum\limits_{k=-b_n+1}^{b_n-1}\left[ \dfrac{1}{n}\left\|\sum_{t=1}^{n-|k|}(X_{st}-\mu)\right\|_{\infty} + \dfrac{1}{n}\left\|\sum_{t=1}^{n-|k|}(X_{s(t+k)}-\mu)\right\|_{\infty} \right]\\
    &\leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \\
    & \quad \times \sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}\left[ \dfrac{1}{n}\left\|\sum_{t=n-|k|+1}^{n}(X_{st} - \mu) - n(\bar{X}_s - \mu) \right\|_{\infty} + \dfrac{1}{n}\left\|\sum_{t=1}^{|k|}(X_{st} - \mu) - n(\bar{X}_s - \mu)\right\|_{\infty} \right]\\
    &\leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \sum_{s=1}^{m}\sum\limits_{k=-b_n+1}^{b_n-1}\left[ \dfrac{1}{n}\left\|\sum_{t=n-|k|+1}^{n}(X_{st} - \mu)\right\|_{\infty} + \dfrac{1}{n}\left\|\sum_{t=1}^{|k|}(X_{st} - \mu)\right\|_{\infty} + 2\|\bar{X}_s - \mu\|_{\infty} \right]\\
    & \leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \sum\limits_{s=1}^{m}\sum_{k=-b_n + 1}^{b_n-1}   \dfrac{1}{n}\left[\left\|\sum_{t=n-|k|+1}^{n}(X_{st} - \mu)\right\|_{\infty} + \left\|\sum_{t=1}^{|k|}(X_{st} - \mu)\right\|_{\infty} \right]\\
    & \; \;+ 2(2b_n - 1)\|\bar{\bar{X}} - \mu\|_{\infty}\|\bar{X}_1 - \mu\|_{\infty} 
\end{align*}
\\
Using SIP on summation of $k$ terms, we obtain the following upper bound for $|M_1|$
\begin{align*}
|M_1|
    & < 2\|(\bar{\bar{X}} - \mu)\|_{\infty}\sum\limits_{k=-b_n + 1}^{b_n-1}\left[ \dfrac{D \psi(k)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{2k \log\log k}}{n}  \right] \\
    & \quad + 2(2b_n - 1)\|\bar{\bar{X}} - \mu\|_{\infty}\|\bar{X}_1 - \mu\|_{\infty} \\
    &\leq 2(2b_n - 1)\|(\bar{\bar{X}} - \mu)\|_{\infty} \left[ \dfrac{D \psi(n)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{n \log\log n}}{n}  \right]\\
    & \quad  + 2(2b_n - 1)\|\bar{\bar{X}} - \mu\|_{\infty}\|\bar{X}_1 - \mu\|_{\infty}\,.
\end{align*}
%
We use Lemma~\ref{lemma:consis_1} to upper bound the above term as
\begin{equation} \label{eq:strong_consis_term-1}
    4(2b_n - 1)\left[ \dfrac{D \psi(n)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{n \log\log n}}{n}  \right]^2
\end{equation}
\\
Using the bound on $\psi(n)$ given by \cite{stra:1964}, equation (\ref{eq:strong_consis_term-1}) becomes $\mathcal{O}\left(b_n \log \log n/n\right)$ which converges to 0 as $n \to \infty$. Now we will show that $M_2 \xrightarrow{a.s.} 0$ as $n\to \infty$

>>>>

In the next section we introduce the new sample autocovariance estimator and calculate its bias for finite samples. We observe that the bias term for R-ACF is similar to the bias of empirical ACF except for a few small order end-effect terms that vanish as $n \to \infty$. In section \ref{sec:variance_est}, we introduce three estimators for $\Sigma$ and study the properties of RSV estimator through proof of strong consistency, bias, and variance calculations. We will elaborate on the estimator for ESS using RSV in section \ref{sec:ess}. Section \ref{sec:examples} gives a simulation study on three different examples for the proposed estimators. There are only a handful of stochastic processes where the true autocovariance and $\Sigma$ are known. To compare the estimators when the truth is known, we use a slowly mixing vector autoregressive process of order 1 (VAR(1)). We successfully show in section \ref{ex:var} that our estimators yield better results as compared to the classical sample ACF and SVE. A more promising advantage of our estimators is observed when the target distribution is multi-modal. For this purpose we use a bivariate bi-modal probability distribution introduced by \cite{gelman1991note} in section \ref{ex:boomerang}. All the important proofs are given in the Appendix.

@@

% \begin{figure}[h]
%     \centering
%     \begin{subfigure}[h]{0.4\textwidth}
%          \centering
%          \includegraphics[width=\textwidth]{plots/scatter_plot_500.pdf}
%          \caption{$n = 500$}
%          \label{subfig:var-sp_500}
%      \end{subfigure}
%      \begin{subfigure}[h]{0.4\textwidth}
%          \centering
%          \includegraphics[width=\textwidth]{plots/scatter_plot_50000.pdf}
%          \caption{$n = 5 \times 10^4$}
%          \label{subfig:var-sp_50000}
%      \end{subfigure}
%     \caption{Scatter plot of five Markov chains for different sample sizes.}
%     \label{fig:var-sp}
% \end{figure}

@@

% \begin{lemma} \label{lemma:consis_1}
%   If Assumption~\ref{ass:sip} holds, then $\| \bar{\bar{X}} - \mu\|_{\infty}, \|\bar{X}_s - \mu\|_{\infty} \xrightarrow[]{a.s.} 0$ as $n \to \infty$ for all $s \in \{1,..., m\}$\\
%   \end{lemma}

% \begin{proof}
% Let $\|.\|$ denote the Euclidean norm. Assumption \ref{ass:sip} allows us to upper bound $\|\bar{X}_s - \mu\|_{\infty}$ using SIP. Following that, we show that the upper bound term converges to 0 as $n \to \infty$ using lemma \ref{lemma: brownian}.
%  \begin{align*}
%     \|\bar{X}_s - \mu\|_{\infty} & \leq \|\bar{X}_s - \mu\| \\
%     &= \dfrac{1}{n}\left\|\sum_{t=1}^{n}X_{st} - n\mu\right\| = \dfrac{1}{n}  \left \|\sum_{t=1}^{n}X_{st} - n\mu \pm L B(n) \right\|\\
%     & \leq \dfrac{1}{n}\left\|\sum_{t=1}^{n}X_{st} - \mu - L B(n)\right\| + \dfrac{\left\|L B(n)\right\|}{n}\\
%     &< \dfrac{D\psi(n)}{n} + \dfrac{\|L B(n)\|}{n}\\
%     &< \dfrac{D\psi(n)}{n} + \dfrac{1}{n}\|L\| \left(\sum\limits_{i=1}^{p}|B^{(i)}(k)|^2\right)^{1/2}\\
%     & \leq \dfrac{D\psi(n)}{n} + \dfrac{1}{n}\|L\| p^{1/2}(1+\epsilon)\sqrt{2n \log\log n}\\
%     & \xrightarrow[]{a.s.} 0\;\; as \;\; n\to \infty
%  \end{align*}
% %
% Similarly one can easily show that $\|\bar{\bar{X}} - \mu\|_{\infty} \to 0$ as $n \to \infty$ using the same upper-bound as:
% % 
% \begin{align*}
%     \|\bar{\bar{X}} - \mu\|_{\infty} & \leq \|\bar{\bar{X}} - \mu\| = \dfrac{1}{m}\left \|\sum_{j=1}^{m}(\bar{X}_j- \mu) \right\| \leq \dfrac{1}{m}\sum_{j=1}^{m}\|\bar{X}_j - \mu\| \xrightarrow{a.s.} 0 \, \textrm{as } n \to \infty 
% \end{align*}
% \end{proof}

@@

\begin{proof}[Proof of Theorem~\ref{th:RAC_expec}]
\begin{align*}
    & \hat{\Gamma}_{RAC,s}(k) \\
   & = \dfrac{1}{n}\sum_{t=1}^{n-|k|} \left(X_{st} - \bar{\bar{X}} \right) \left(X_{s(t+k)} - \bar{\bar{X}} \right)^T \\
    &= \left[\dfrac{1}{n} \sum_{t=1}^{n-|k|} \left(X_{st} - \bar{X}_s \right) \left(X_{s(t+k)} - \bar{X}_s \right)^T \right] + \left[\dfrac{1}{n} \sum_{t=1}^{|k|} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - X_{st} \right)^T\right] \\ 
    & \quad +  \left[\dfrac{1}{n} \sum_{t=n-|k|+1}^{n}  \left( \bar{X}_s - X_{st} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right] + \left[\dfrac{n-|k|}{n} \left(\bar{X}_s - \bar{\bar{X}}\right)   \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \hat{\Gamma}_s(k) - \dfrac{1}{n} \sum_{t=1}^{|k|}A_{st}^T - \dfrac{1}{n}\sum_{t=n-|k|+1}^{n}A_{st} + \dfrac{n-|k|}{n}  \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,, \numberthis \label{eq:acf_breakdown}
\end{align*}
%
where $A_{st} = (X_{st}-\bar{X}_s)(\bar{X}_s - \bar{\bar{X}})^T$. We will study the expectations of the each of the above terms. Without loss of generality, consider $A_{11}$,
\begin{align*}
    &\E \left[A_{11} \right]\\
     &= \mathbb{E} \left[ \left(X_{11} - \bar{X}_1 \right) \left(\bar{X}_1 - \bar{\bar{X}} \right)^T \right]\\
     &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E}\left[X_{11}\bar{\bar{X}}^T \right] + \mathbb{E} \left[\bar{X}_1\bar{\bar{X}}^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{1}{m}\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \dfrac{m-1}{m}\mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \dfrac{1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right] + \dfrac{m-1}{m}\mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\\
    &= \dfrac{m-1}{m}\left(\mathbb{E} \left[X_{11}\bar{X}_1^T \right] - \mathbb{E} \left[X_{11}\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\bar{X}_2^T \right] - \mathbb{E} \left[\bar{X}_1\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{t=1}^{n}\mathbb{E} \left[X_{11}X_{1t}^T \right] - \mathbb{E}\left[X_{11}\right] \mathbb{E}\left[\bar{X}_2^T \right] + \mathbb{E} \left[\bar{X}_1\right]\mathbb{E}\left[\bar{X}_2^T\right] - \text{Var}\left[\bar{X}_1 \right] - \mathbb{E}\left[\bar{X}_1 \right]\mathbb{E}\left[\bar{X}_1^T \right]\right)\\
    &= \dfrac{m-1}{m}\left(\dfrac{1}{n}\sum_{k=0}^{n-1}\Gamma(k) + \mu \mu^T - \mu \mu^T + \mu \mu^T - \text{Var}\left[\bar{X}_1 \right] - \mu \mu^T\right)\\
    &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - n\text{Var}\left[\bar{X}_1 \right]\right)\,. \numberthis \label{eq:acf_2}
\end{align*}
%
Similarly,
%    
\begin{align}
\label{eq:acf_3}
    \mathbb{E} \left[ A_{11}^T \right] &= \mathbb{E}\left[ A_{11}\right]^T = \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k)^T - n\text{Var}\left[\bar{X}_1 \right] \right)\,.
    % &= \dfrac{m-1}{mn}\left(\sum_{k=0}^{n-1}\Gamma(k) - \Sigma \right)\\
\end{align}

Further,
\begin{align*}
\mathbb{E}\left[ \left(\bar{X}_{s} - \bar{\bar{X}} \right)  \left(\bar{X}_{s} - \bar{\bar{X}} \right)^{T}\right] &= \mathbb{E}\left[ \bar{X}_{s}\bar{X}_{s}^{T} - \bar{X}_{s}\bar{\bar{X}}^{T} - \bar{\bar{X}} \bar{X}_{s}^{T} + \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= \mathbb{E}\left[ \bar{X}_{s}\bar{X}_{s}^{T} - \bar{\bar{X}}\bar{\bar{X}}^{T}\right]\\
&= \left(\mathbb{E}\left[\bar{X}_{1}\bar{X}_{1}^{T}\right] - \mathbb{E}\left[\bar{\bar{X}}\bar{\bar{X}}^{T}\right]\right)\\
&= \left(\Var(\bar{X}_{1}) + \mu\mu^{T} - \Var(\bar{\bar{X}}) - \mu\mu^{T}\right)\\
&= \dfrac{m-1}{m}\Var(\bar{X}_1)\,. \numberthis \label{eq:acf_4}
\end{align*}
%
Using \eqref{eq:priestly}, \eqref{eq:acf_2}, \eqref{eq:acf_3}, and \eqref{eq:acf_4} in \eqref{eq:acf_breakdown},
\begin{align*}
    & \E \left[\hat{\Gamma}_{RAC,s}(k) \right] \\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - \dfrac{1}{n} \left(\sum\limits_{t=1}^{|k|}\mathbb{E}[A_{1t}^T] + \sum\limits_{t=n-|k|+1}^{n}\mathbb{E}[A_{1t}]\right) + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \mathbb{E}\left[\hat{\Gamma}_{s}(k)\right] - {\color{red}\dfrac{|k|}{n}\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h) + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T - 2 \Var(\bar{X}_1)\right)} + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)\\
    &= \left(1- \dfrac{|k|}{n}\right)\Gamma(k) - \dfrac{|k|}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)\right) - \left(2-\dfrac{1}{m}\right) \Var(\bar{X}_1)\right] - \dfrac{\Var(\overline{X}_1)}{m}\,.
\end{align*}
%
 We can use the results of \cite{song1995optimal} given in \eqref{eq:song} to expand $\Var(\bar{X}_1)$. Expectation of $\hat{\Gamma}_{RAC,s}(k)$ can then broken down as following,
 \begin{equation} \label{eq:rac_expec_breakdown}
     \mathbb{E}\left[\hat{\Gamma}_{RAC}(k)\right] = \left(1- \dfrac{|k|}{n}\right)\Gamma(k) + O_1 + O_2\,.
 \end{equation}
%
where,
\begin{align*}
    O_1 &= -\dfrac{|k|}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma(h)\right) - \left(2-\dfrac{1}{m}\right) \left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right)\right] \, ,\\
    O_2 &= -\dfrac{1}{m}\left(\dfrac{\Sigma}{n} + \dfrac{\Phi}{n^2}\right) + o(n^{-2})
\end{align*}
%
We observe that both $O_1 \textrm{ and } O_2$ are small order terms that converge to 0 as $n \to \infty$. Here, $O_1 = (-\abs{k}/n)\mathcal{O}(1/n) \textrm{ and } O_2 = \mathcal{O}(1/n)$. For a diagonal element of $\Gamma$,
% We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.
%
\begin{align*}
& \E \left[ \hat{\Gamma}_{RAC,s}^{ii}\right]\\
& = \mathbb{E}\left[\hat{\Gamma}^{ii}_{s}(k)\right] - \dfrac{|k|}{n}\left[\left(1-\dfrac{1}{m}\right)\left(\dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma^{ii}(h)^T + \dfrac{1}{n}\sum_{h=0}^{n-1}\Gamma^{ii}(h)\right) - \left(2-\dfrac{1}{m}\right) \Var(\overline{X}_1)^{ii}\right] - \dfrac{\Var(\overline{X}_1)^{ii}}{m}\,.
\end{align*}
In the presence of positive correlation, the leftover term is positive.
\end{proof}

@@

Consequently
\left|\hat{\Sigma}_{RSV}^{ij} - \tilde{\Sigma}^{ij}  \right| = |M_1 + M_2| \leq |M_1| + |M_2|\,.
We first present a result which will be useful to use later. For any Markov chain $s$, 
  \|\bar{X}_s - \mu\|_{\infty} & \leq \|\bar{X}_s - \mu\| = \dfrac{1}{mn}\left\|\sum_{t=1}^{n}X_{st} - n\mu\right\| \\
  &= \dfrac{1}{n}  \left \|\sum_{t=1}^{n}X_{st} - n\mu \pm L B(n) \right\|\\
  & \leq \dfrac{1}{n}\left\|\sum_{t=1}^{n}X_{st} - n \mu - L B(n)\right\| + \dfrac{\left\|L B(n)\right\|}{n}\\
  &< \dfrac{D\psi(n)}{n} + \dfrac{\|L B(n)\|}{n}\\
  &< \dfrac{D\psi(n)}{n} + \dfrac{1}{n}\|L\| \left(\sum\limits_{i=1}^{p}|B^{(i)}(k)|^2\right)^{1/2}\\
  & \leq \dfrac{D\psi(n)}{n} + \dfrac{1}{n}\|L\| p^{1/2}(1+\epsilon)\sqrt{2n \log\log n}\,. \numberthis \label{eq:xbars_bound}
  % & \xrightarrow[]{a.s.} 0\;\; as \;\; n\to \infty 
 Similarly,
\label{eq:xbarbar_bound}
   \| \bar{\bar{X}} - \mu\|_{\infty} \leq \dfrac{D\psi(n)}{n} + \dfrac{1}{n}\|L\| p^{1/2}(1+\epsilon)\sqrt{2n \log\log n}\,.
